---
title: "Plots - artificial vs. natural experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis_all.cycles/derived_data/"
dir_out <- "analysis_all.cycles/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---

# Load packages
```{r Libraries}
pack_to_load <- c("R.utils", "tools", "ggplot2", "doBy", "tidyverse", "patchwork", "ggsci")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)
```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```



# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)


# add another column which concatenates Sample and Spot 
imp_data$ID <- paste(imp_data$Sample, imp_data$Spot)
imp_data <- imp_data[c(1:3, 44, 4:43)]

# copy the column Cycle and convert into numeric
imp_data$Stroke <- paste(imp_data$Cycle) 
imp_data[["Stroke"]] <- as.numeric(imp_data[["Stroke"]],levels=c("0", "50", "250", "1000",
                        "2000"))

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 10:44

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```


---


# Plot each of the selected numeric variable 
## Sorted by contact material 
### artificial vs. natural bone: flint samples 
```{r}

flint <- filter(imp_data, Raw.material == "flint")


for (i in num.var){
 
 # get the min/max range of the data set
 range_var <- range(flint[i]) 
  
    # plot bone plate
    f_bp <- ggplot(flint[grep("bone plate", flint[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) + 
            labs(title = "flint - bone plate", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            geom_line(aes(group = Spot,), colour = "#FAD510") +  
            scale_colour_manual(values = "#FAD510") +
            theme_classic() 

    # plot cow scapula 
    f_cs <- ggplot(flint[grep("cow scapula", flint[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "flint - cow scapula", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            scale_colour_manual(values = "#CB2314") +
            geom_line(aes(group = Spot,), colour = "#CB2314") +
            theme_classic() 

	  # combine the flint plots 
    f.bone <- f_bp + f_cs + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
       
    print(f.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_f.bone_", names(imp_data)[i],
	            ".pdf")
	ggsave(filename = file_out, plot = f.bone, path = dir_out, device = "pdf")

}

```


### artificial vs. natural bone: lydite samples 
```{r}

lydite <- filter(imp_data, Raw.material == "lydite")

for (i in num.var){
	
  # get the min/max range of the data set
  range_var <- range(lydite[i]) 
  
    # plot bone plate 
    l_bp <- ggplot(lydite[grep("bone plate", lydite[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "lydite - bone plate", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            geom_line(aes(group = Spot,), colour = "#FAD510") +  
            scale_colour_manual(values = "#FAD510") +
            theme_classic() 
    
    # plot cow scapula 
    l_cs <- ggplot(lydite[grep("cow scapula", lydite[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "lydite - cow scapula", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            scale_colour_manual(values = "#CB2314") +
            geom_line(aes(group = Spot,), colour = "#CB2314") +
            theme_classic() 

	  # combine the lydite plots 
    l.bone <- l_bp + l_cs + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
   
     print(l.bone)
 
  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_l.bone_", 
	            names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = l.bone, path = dir_out, device = "pdf")
	

}

```
 
 
### artificial vs. natural skin: flint samples 
```{r}

for (i in num.var){   
	
  # get the min/max range of the data set
  range_var <- range(flint[i]) 
  
    # plot skin pad 
    f_sp <- ggplot(flint[grep("skin pad", flint[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "flint - skin pad", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            geom_line(aes(group = Spot,), colour = "#046C9A") + 
            scale_colour_manual(values = "#046C9A") +
            theme_classic()  

    # plot pig skin 
    f_ps <- ggplot(flint[grep("pig skin", flint[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "flint - pig skin", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            scale_colour_manual(values = "#52854c") +
            geom_line(aes(group = Spot,), colour = "#52854c") +
            theme_classic() 

	  # combine the flint and the lydite plots 
    f.skin <- f_sp + f_ps + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
       
    print(f.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_f.skin_", 
	            names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = f.skin, path = dir_out, device = "pdf")

}

```


### artificial vs. natural skin: lydite samples 
```{r}

for (i in num.var){
	
  # get the min/max range of the data set
  range_var <- range(lydite[i]) 
  
    # plot skin pad 
    l_sp <- ggplot(lydite[grep("skin pad", lydite[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) + 
            labs(title = "lydite - skin pad", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            geom_line(aes(group = Spot,), colour = "#046C9A") + 
            scale_colour_manual(values = "#046C9A") +
            theme_classic()

    # plot pig skin 
    l_ps <- ggplot(lydite[grep("pig skin", lydite[["Contact.material"]]), ], 
            aes_string(y = names(imp_data)[i], x = "Cycle")) +
            geom_point(aes(shape = Spot, colour = Sample), size = 3) +
            labs(title = "lydite - pig skin", y = names(imp_data)[i], x = "cycle") + 
            coord_cartesian(ylim = range_var) +
            guides(colour = "none") +
            scale_colour_manual(values = "#52854c") +
            geom_line(aes(group = Spot,), colour = "#52854c") +
            theme_classic()

	  # combine the flint and the lydite plots 
    l.skin <- l_sp + l_ps + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
    print(l.skin)


  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_l.skin_", names(imp_data)[i],
	            ".pdf")
	ggsave(filename = file_out, plot = l.skin, path = dir_out, device = "pdf")


}

```



# Plot the mean value (3 measurements) of each of the selected numeric variable 
## Sorted by contact material 
### cow scapula & bone plate: Sq
```{r}
# remove possible outliers (based on values and prior plots)
imp_data2 <- imp_data[-c(1, 4, 7, 10, 13, 18, 32, 47, 75), ]

# add a column that combines sample and cycle
ID.cycle <- unite(imp_data2, ID_Cycle, c(Sample, Cycle), remove = FALSE)

# compute the mean of the three spots per sample 
mean_cycle <- summaryBy(.~ ID_Cycle + Sample + Stroke + Raw.material + Contact.material, 
                         data = ID.cycle, FUN = mean)

# define new num.var for mean_cycle 
num.var2 <- 6:length(mean_cycle)
for (j in num.var2) cat("[",j,"] ", names(mean_cycle)[j], "\n", sep="")

     # plot bone plate 
     Sq.bp <- ggplot(mean_cycle[grep("bone plate", mean_cycle[["Contact.material"]]), ], 
              aes_string(y = "Sq.mean", x = "Stroke")) +
              geom_point(aes(colour = Sample), size = 3) +
              geom_line(aes(group = Sample)) +  
              scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
              labs(title = "bone plate", y = NULL, x = "cycle") + 
              coord_cartesian(ylim = c(0, 7000)) + 
              scale_colour_manual(values = c("#FAD50F", "#FAD50F")) +
              theme_classic() 
    
    # plot cow scapula 
    Sq.cs <- ggplot(mean_cycle[grep("cow scapula", mean_cycle[["Contact.material"]]), ],
             aes_string(y = "Sq.mean", x = "Stroke")) +
             geom_point(aes(colour = Sample), size = 3) +
             geom_line(aes(group = Sample)) +
             scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
             coord_cartesian(ylim = c(0, 7000)) +
             labs(title = "cow scapula", y = "ΔSq [nm]", x = "cycle") + 
             scale_colour_manual(values = c("#CB2213", "#CB2213")) +
             theme_classic() 


 	 # combine the plots 
   Sq.bone.mean <- Sq.cs + Sq.bp + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
       
   print(Sq.bone.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Sq.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Sq.bone.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")


```


### pig skin & skin pad: Sq
```{r}
  Sq.sp <- ggplot(mean_cycle[grep("skin pad", mean_cycle[["Contact.material"]]), ],
           aes_string(y = "Sq.mean", x = "Stroke")) +
           geom_point(aes(colour = Sample), size = 3) +
           geom_line(aes(group = Sample)) +
           scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
           labs(title = "skin pad", y = NULL, x = "cycle") + 
           coord_cartesian(ylim = c(0, 6000)) +  
           scale_colour_manual(values = c("#52854B", "#52854B")) +
           theme_classic() 

    Sq.ps <- ggplot(mean_cycle[grep("pig skin", mean_cycle[["Contact.material"]]), ],
             aes_string(y = "Sq.mean", x = "Stroke")) +
             geom_point(aes(colour = Sample), size = 3) +
             geom_line(aes(group = Sample)) +
             scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
             coord_cartesian(ylim = c(0, 6000)) +
             labs(title = "pig skin", y = "ΔSq [nm]", x = "cycle") + 
             scale_colour_manual(values = c("#036C9A", "#036C9A")) +
             theme_classic() 

	  # combine the plots 
    Sq.skin.mean <- Sq.ps + Sq.sp + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
       
    print(Sq.skin.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Sq.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Sq.skin.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")

```


### cow scapula & bone plate: Vmc
```{r}

     Vmc.bp <- ggplot(mean_cycle[grep("bone plate", mean_cycle[["Contact.material"]]), ],
               aes_string(y = "Vmc.mean", x = "Stroke")) +
               geom_point(aes(colour = Sample), size = 3) +
               geom_line(aes(group = Sample)) +
               scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +                 
               labs(title = "bone plate", y = NULL, x = "cycle") + 
               coord_cartesian(ylim = c(0, 3.5)) + 
               scale_colour_manual(values = c("#FAD50F", "#FAD50F")) +
               theme_classic() 

     Vmc.cs <- ggplot(mean_cycle[grep("cow scapula", mean_cycle[["Contact.material"]]), ],
              aes_string(y = "Vmc.mean", x = "Stroke")) +
              geom_point(aes(colour = Sample), size = 3) +
              geom_line(aes(group = Sample)) +
              scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
              coord_cartesian(ylim = c(0, 3.5)) +
              labs(title = "cow scapula", y = "Vmc [µm³/µm²]", x = "cycle") + 
              scale_colour_manual(values = c("#CB2213", "#CB2213")) +
              theme_classic() 

     # combine the plots 
     Vmc.bone.mean <- Vmc.cs + Vmc.bp + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
     print(Vmc.bone.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Vmc.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Vmc.bone.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")


```


### pig skin & skin pad: Vmc
```{r}
  Vmc.sp <- ggplot(mean_cycle[grep("skin pad", mean_cycle[["Contact.material"]]), ],
            aes_string(y = "Vmc.mean", x = "Stroke")) +
            geom_point(aes(colour = Sample), size = 3) +
            geom_line(aes(group = Sample)) +
            scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
            labs(title = "skin pad", y = NULL, x = "cycle") + 
            coord_cartesian(ylim = c(0, 3.5)) +  
            scale_colour_manual(values = c("#52854B", "#52854B")) + 
            theme_classic() 

    Vmc.ps <- ggplot(mean_cycle[grep("pig skin", mean_cycle[["Contact.material"]]), ],
              aes_string(y = "Vmc.mean", x = "Stroke")) +
              geom_point(aes(colour = Sample), size = 3) +
              geom_line(aes(group = Sample)) +
              scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
              coord_cartesian(ylim = c(0, 3.5)) +
              labs(title = "pig skin", y = "Vmc [µm³/µm²]", x = "cycle") + 
              scale_colour_manual(values = c("#036C9A", "#036C9A")) +
              theme_classic() 

  	# combine the plots 
    Vmc.skin.mean <- Vmc.ps + Vmc.sp + plot_layout(width = c(3/6, 3/6), guides = 'collect')  
    print(Vmc.skin.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Vmc.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Vmc.skin.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")

```


### cow scapula & bone plate: HAsfc9
```{r}

     HAsfc9.bp <- ggplot(mean_cycle[grep("bone plate", mean_cycle[["Contact.material"]]), ],
                  aes_string(y = "HAsfc9.mean", x = "Stroke")) +
                  geom_point(aes(colour = Sample), size = 3) +
                  geom_line(aes(group = Sample)) +
                  scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                  labs(title = "bone plate", y = NULL, x = "cycle") + 
                  coord_cartesian(ylim = c(0.0, 11.0)) +
                  scale_colour_manual(values = c("#FAD50F", "#FAD50F")) +
                 theme_classic() 

    HAsfc9.cs <- ggplot(mean_cycle[grep("cow scapula", mean_cycle[["Contact.material"]]), ],
                 aes_string(y = "HAsfc9.mean", x = "Stroke")) +
                 geom_point(aes(colour = Sample), size = 3) +
                 geom_line(aes(group = Sample)) +
                 scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                 coord_cartesian(ylim = c(0.0, 11.0)) +
                 labs(title = "cow scapula", y = "HAsfc9", x = "cycle") + 
                 scale_colour_manual(values = c("#CB2213", "#CB2213")) +
                 theme_classic() 

	  # combine the plots 
    HAsfc9.bone.mean <- HAsfc9.cs + HAsfc9.bp + plot_layout(width = c(3/6, 3/6), 
                        guides = 'collect')  
       
    print(HAsfc9.bone.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_HAsfc9.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = HAsfc9.bone.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")


```


### pig skin & skin pad: HAsfc9
```{r}
  HAsfc9.sp <- ggplot(mean_cycle[grep("skin pad", mean_cycle[["Contact.material"]]), ],
               aes_string(y = "HAsfc9.mean", x = "Stroke")) +
               geom_point(aes(colour = Sample), size = 3) +
               geom_line(aes(group = Sample)) +
               scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
               labs(title = "skin pad", y = NULL, x = "cycle") + 
               coord_cartesian(ylim = c(0.0, 11.0)) +  
               scale_colour_manual(values = c("#52854B", "#52854B")) +
               theme_classic() 


    HAsfc9.ps <- ggplot(mean_cycle[grep("pig skin", mean_cycle[["Contact.material"]]), ],
                 aes_string(y = "HAsfc9.mean", x = "Stroke")) +
                 geom_point(aes(colour = Sample), size = 3) +
                 geom_line(aes(group = Sample)) +
                 scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                 coord_cartesian(ylim = c(0.0, 11.0)) +
                 labs(title = "pig skin", y = "HAsfc9", x = "cycle") + 
                 scale_colour_manual(values = c("#036C9A", "#036C9A")) +
                theme_classic() 

  	# combine the plots 
    HAsfc9.skin.mean <- HAsfc9.ps + HAsfc9.sp + plot_layout(width = c(3/6, 3/6), guides =
                      'collect')  
       
    print(HAsfc9.skin.mean)
  
  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_HAsfc9.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = HAsfc9.skin.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")

```


### cow scapula & bone plate: epLsar
```{r}

     epLsar.bp <- ggplot(mean_cycle[grep("bone plate", mean_cycle[["Contact.material"]]), ],
                  aes_string(y = "epLsar.mean", x = "Stroke")) +
                  geom_point(aes(colour = Sample), size = 3) +
                  geom_line(aes(group = Sample)) +
                  scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                  labs(title = "bone plate", y = NULL, x = "cycle") + 
                  coord_cartesian(ylim = c(0.000, 0.005)) +
                  scale_colour_manual(values = c("#FAD50F", "#FAD50F")) +
                 theme_classic() 

    epLsar.cs <- ggplot(mean_cycle[grep("cow scapula", mean_cycle[["Contact.material"]]), ],
                 aes_string(y = "epLsar.mean", x = "Stroke")) +
                 geom_point(aes(colour = Sample), size = 3) +
                 geom_line(aes(group = Sample)) +
                 scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                 coord_cartesian(ylim = c(0.000, 0.005)) +
                 labs(title = "cow scapula", y = "epLsar", x = "cycle") + 
                 scale_colour_manual(values = c("#CB2213", "#CB2213")) +
                 theme_classic() 

	  # combine the plots 
    epLsar.bone.mean <- epLsar.cs + epLsar.bp + plot_layout(width = c(3/6, 3/6), 
                        guides = 'collect')  
       
    print(epLsar.bone.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_epLsar.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = epLsar.bone.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")


```


### pig skin & skin pad: epLsar
```{r}
  epLsar.sp <- ggplot(mean_cycle[grep("skin pad", mean_cycle[["Contact.material"]]), ],
               aes_string(y = "epLsar.mean", x = "Stroke")) +
               geom_point(aes(colour = Sample), size = 3) +
               geom_line(aes(group = Sample)) +
               scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
               labs(title = "skin pad", y = NULL, x = "cycle") + 
               coord_cartesian(ylim = c(0.000, 0.005)) +  
               scale_colour_manual(values = c("#52854B", "#52854B")) +
               theme_classic() 

    epLsar.ps <- ggplot(mean_cycle[grep("pig skin", mean_cycle[["Contact.material"]]), ],
                 aes_string(y = "epLsar.mean", x = "Stroke")) +
                 geom_point(aes(colour = Sample), size = 3) +
                 geom_line(aes(group = Sample)) +
                 scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                 coord_cartesian(ylim = c(0.000, 0.005)) +
                 labs(title = "pig skin", y = "epLsar", x = "cycle") + 
                 scale_colour_manual(values = c("#036C9A", "#036C9A")) +
                 theme_classic() 

	  # combine the plots 
    epLsar.skin.mean <- epLsar.ps + epLsar.sp + plot_layout(width = c(3/6, 3/6), 
                        guides = 'collect')  
       
  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_epLsar.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = epLsar.skin.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")

```


### cow scapula & bone plate: Asfc
```{r}

     Asfc.bp <- ggplot(mean_cycle[grep("bone plate", mean_cycle[["Contact.material"]]), ],
                aes_string(y = "Asfc.mean", x = "Stroke")) +
                geom_point(aes(colour = Sample), size = 3) +
                geom_line(aes(group = Sample)) +
                scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
                labs(title = "bone plate", y = NULL, x = "cycle") + 
                coord_cartesian(ylim = c(0, 35)) +
                scale_colour_manual(values = c("#FAD50F", "#FAD50F")) +
                theme_classic() 

    Asfc.cs <- ggplot(mean_cycle[grep("cow scapula", mean_cycle[["Contact.material"]]), ],
               aes_string(y = "Asfc.mean", x = "Stroke")) +
               geom_point(aes(colour = Sample), size = 3) +
               geom_line(aes(group = Sample)) +
               scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
               coord_cartesian(ylim = c(0, 35)) +
               labs(title = "cow scapula", y = "Asfc", x = "cycle") + 
               scale_colour_manual(values = c("#CB2213", "#CB2213")) +
               theme_classic() 

  	# combine the plots 
    Asfc.bone.mean <- Asfc.cs + Asfc.bp + plot_layout(width = c(3/6, 3/6), 
                      guides = 'collect')  
       
    print(Asfc.bone.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Asfc.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Asfc.bone.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")


```


### cow scapula & bone plate: Asfc
```{r}
  Asfc.sp <- ggplot(mean_cycle[grep("skin pad", mean_cycle[["Contact.material"]]), ],
             aes_string(y = "Asfc.mean", x = "Stroke")) +
             geom_point(aes(colour = Sample), size = 3) +
             geom_line(aes(group = Sample)) +
             scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) +
             labs(title = "skin pad", y = NULL, x = "cycle") + 
             coord_cartesian(ylim = c(0, 35)) + 
             scale_colour_manual(values = c("#52854B", "#52854B")) +
             theme_classic() 

    Asfc.ps <- ggplot(mean_cycle[grep("pig skin", mean_cycle[["Contact.material"]]), ],
               aes_string(y = "Asfc.mean", x = "Stroke")) +
               geom_point(aes(colour = Sample), size = 3) +
               geom_line(aes(group = Sample)) +
               scale_x_continuous(breaks = c(0, 50, 250, 1000, 2000)) + 
               coord_cartesian(ylim = c(0, 35)) +
               labs(title = "pig skin", y = "Asfc", x = "cycle") + 
               scale_colour_manual(values = c("#036C9A", "#036C9A")) +
               theme_classic() 

	  # combine the plots 
    Asfc.skin.mean <- Asfc.ps + Asfc.sp + plot_layout(width = c(3/6, 3/6), 
                      guides = 'collect')  
       
    print(Asfc.skin.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Asfc.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Asfc.skin.mean, path = dir_out, device = "pdf", 
	       width = 350, height = 270, units = "mm")

```



---


# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis_all.cycles/scripts/RStudioVersion.txt", n = 1)`.


---

# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)
```



---

END OF SCRIPT
