---
title: "Plots - artificial vs. natural experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
---

---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  



```{r}
dir_in <- "analysis_before.after/derived_data/"
dir_out <- "analysis_before.after/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r Libraries}
pack_to_load <- c("R.utils", "tools", "ggplot2", "doBy", "tidyverse", "patchwork", 
                "ggsci", "ggfortify")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)
```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, 
           row.names = NULL)

```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 9:length(imp_data)

```


The following variables will be used: 


```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```


---


# Calculate and plot the absolute difference for each sample and each parameter 
## Sorted by contact material 
### pig skin & skin pad  
```{r}
# add another coloum which concatenates Sample and Spot 
imp_data$ID <- paste(imp_data$Sample, imp_data$Spot)

diff_data <- imp_data[-(6:8)] %>% 
  pivot_wider(names_from = Cycle, values_from = all_of(names(imp_data)[num.var]))


output_list <- vector(mode = "list", length = length(names(imp_data)[num.var]))
names(output_list) <- paste0(names(imp_data)[num.var], ".diff")

for (i in seq_along(num.var)){
  temp <- select(diff_data, starts_with(paste0(names(imp_data[num.var])[i], "_")))
  output_list[[i]] <- unlist(select(temp, contains("2000")) - select(temp, 
                      contains("before")))
}

var.diff <- as.data.frame(output_list)

# create a data frame with the differences from before and after as values 
difference <- as.data.frame(cbind(select(diff_data, Sample:ID), var.diff))
row.names(difference) <- NULL


# define numeric variable 
num.var2 <- 6:length(difference)
# the following variables will be used: 
for (j in num.var2) cat("[",j,"] ", names(difference)[j], "\n", sep="")


for (j in num.var2){
  
  # get the min/max range of the data set
  range_var <- range(difference[j])  
  # plot pig skin 
  p_dif_skin <- ggplot(data = difference[grep("pig skin", difference[["Contact.material"]]),
                ], aes_string(x = "Spot", names(difference)[j], colour = "Sample")) + 
	              geom_jitter(size = 3, width = 0.25) +
                theme(legend.position = "none") +
                coord_cartesian(ylim = range_var) +
	              facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                labs(title = "natural contact material") + 
                ylab(names(difference)[j]) + xlab(NULL) +
	              labs(y = gsub("\\.", " ", names(difference)[j])) +
	              scale_colour_futurama() +
	              theme_classic()

   # plot skin pad 
   p_dif_skin.pad <- ggplot(data = difference[grep("skin pad",
                     difference[["Contact.material"]]), ], aes_string(x = "Spot",    
                     names(difference)[j], 
	                   colour = "Sample")) + 
	                   geom_jitter(size = 3, width = 0.25) +
                     theme(legend.position = "none") +
                     coord_cartesian(ylim = range_var) +
	                   facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                     labs(title = "artificial contact material") + 
                     ylab(names(difference)[j]) + xlab(NULL) +
	                   labs(y = gsub("\\.", " ", names(difference)[j])) +
	                   scale_colour_futurama() +
	                   theme_classic()

	# combine both plots 
  p_dif.skin <- p_dif_skin + p_dif_skin.pad + plot_layout(width = c(3/6, 3/6), 
                guides = 'collect')  
  print(p_dif.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_skin_", names(difference)[j],
	            ".pdf")
	ggsave(filename = file_out, plot = p_dif.skin, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")

}

```

# cow scapula & bone plate 
```{r}
for (j in num.var2){
 # get the min/max range of the data set
 range_var <- range(difference[j])  
  
  # plot cow scapula 
  p_dif_bone <- ggplot(data = difference[grep("cow scapula",   
                difference[["Contact.material"]]), ], aes_string(x = "Spot", 
                names(difference)[j], colour = "Sample")) + 
	              geom_jitter(size = 3, width = 0.25) +
                theme(legend.position = "none") +
                coord_cartesian(ylim = range_var) +
	              facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                labs(title = "natural contact material") + 
                ylab(names(difference)[j]) + xlab(NULL) +
	              labs(y = gsub("\\.", " ", names(difference)[j])) +
	              scale_colour_futurama() +
	              theme_classic()
 
   # plot bone plate 
   p_dif_bone.plate <- ggplot(data = difference[grep("bone plate", 
                       difference[["Contact.material"]]), ], aes_string(x = "Spot", 
                       names(difference)[j], colour = "Sample")) + 
	                     geom_jitter(size = 3, width = 0.25) +
                       theme(legend.position = "none") +
                       coord_cartesian(ylim = range_var) +
	                     facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                       labs(title = "artificial contact material") + 
                       ylab(names(difference)[j]) + xlab(NULL) +
	                     labs(y = gsub("\\.", " ", names(difference)[j])) +
	                     scale_colour_futurama() +
	                     theme_classic()
 
	# combine both plots 
  p_dif.bone <- p_dif_bone + p_dif_bone.plate + plot_layout(width = c(3/6, 3/6),
                guides = 'collect')  
  print(p_dif.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_bone_", names(difference)[j],
	            ".pdf")
	ggsave(filename = file_out, plot = p_dif.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")

}
```


# Calculate and plot the absolute difference for each sample and each parameter as mean value
## sorted by contact material 
### cow scapula & bone plate 
```{r}
# compute the mean of the three spots per sample 
mean.sample <- summaryBy(.~ Sample + Raw.material + Contact.material, 
               data = difference, FUN = mean)
# define new num.var for mean_cycle 
num.var3 <- 4:length(mean.sample)


# the following variables will be used: 
for (k in num.var3) cat("[",k,"] ", names(difference)[k], "\n", sep="")

 
for (k in num.var3){
  # get the min/max range of the data set
  range_var <- range(mean.sample[k])  
  # plot cow scapula
  bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
               mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
               names(mean.sample)[k], colour = "Sample")) + 
               geom_point(size = 4) +
               coord_cartesian(ylim = range_var) +
	             facet_wrap(Sample ~ Contact.material, nrow = 2) + 
               labs(title = "natural contact material") + 
               ylab(names(mean.sample)[j]) + xlab(NULL) +
	             labs(y = gsub("\\.", " ", names(mean.sample)[k])) +
	             scale_colour_futurama() +
	             theme_classic() +
               theme(legend.position = "none") +
               theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot bone plate 
   plate.mean <- ggplot(data = mean.sample[grep("bone plate",
                 mean.sample[["Contact.material"]]), ], aes_string(x = "Sample",
                 names(mean.sample)[k], colour = "Sample")) + 
	               geom_point(size = 4) +
                 coord_cartesian(ylim = range_var) +
	               facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                 labs(title = "artificial contact material") + 
                 ylab(names(mean.sample)[j]) + xlab(NULL) +
	               labs(y = gsub("\\.", " ", names(mean.sample)[k])) +
	               scale_colour_futurama() +
	               theme_classic() +
                 theme(legend.position = "none") +
                 theme(axis.text.x = element_blank(),axis.ticks = element_blank())  


	# combine both plots 
  bone.plate.mean <- bone.mean + plate.mean + plot_layout(width = c(3/6, 3/6),
                     guides = 'collect')  
  print(bone.plate.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_bone_", names(mean.sample)[k],
	            ".pdf")
	ggsave(filename = file_out, plot = bone.plate.mean, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")

} 
 

```


### pig skin & skin pad  
```{r}

  # plot skin
for (k in num.var3){
  # get the min/max range of the data set
  range_var <- range(mean.sample[k])  

  skin.mean <- ggplot(data = mean.sample[grep("pig skin", 
               mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
               names(mean.sample)[k], colour = "Sample")) + 
               geom_point(size = 4) +
               coord_cartesian(ylim = range_var) +
	             facet_wrap(Sample ~ Contact.material, nrow = 2) + 
               labs(title = "natural contact material") + 
               ylab(names(mean.sample)[j]) + xlab(NULL) +
	             labs(y = gsub("\\.", " ", names(mean.sample)[k])) +
	             scale_colour_futurama() +
	             theme_classic() +
               theme(legend.position = "none") +
               theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
   # plot skin pad 
   pad.mean <- ggplot(data = mean.sample[grep("skin pad",
               mean.sample[["Contact.material"]]), ], aes_string(x = "Sample",
               names(mean.sample)[k], colour = "Sample")) + 
	             geom_point(size = 4) +
               coord_cartesian(ylim = range_var) +
	             facet_wrap(Sample ~ Contact.material, nrow = 2) + 
               labs(title = "artificial contact material") + 
               ylab(names(mean.sample)[j]) + xlab(NULL) +
	             labs(y = gsub("\\.", " ", names(mean.sample)[k])) +
	             scale_colour_futurama() +
	             theme_classic() +
               theme(legend.position = "none") +
               theme(axis.text.x = element_blank(),axis.ticks = element_blank())  


	# combine both plots 
  skin.pad.mean <- skin.mean + pad.mean + plot_layout(width = c(3/6, 3/6), 
                   guides = 'collect')  
  print(skin.pad.mean)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_skin_", names(mean.sample)[k],
	            ".pdf")
	ggsave(filename = file_out, plot = skin.pad.mean, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")

} 
 

```


## cow scapula & bone plate (with specific range per parameter) 
### Sq
```{r}
  # plot cow scapula 
  Sq_bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
                  mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                  y = "Sq.diff.mean", colour = "Sample")) + 
                  geom_point(size = 4) +
                  coord_cartesian(ylim = c(-2100, 3700)) +
	                facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                  labs(title = "natural contact material") + 
                  ylab("Sq [nm]") + xlab(NULL) +
	                scale_colour_futurama() +
	                theme_classic() +
                  theme(legend.position = "none") +
                  theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
   # plot bone plate 
   Sq.plate.mean <- ggplot(data = mean.sample[grep("bone plate",
                    mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                    y = "Sq.diff.mean", colour = "Sample")) + 
	                  geom_point(size = 4) +
                    coord_cartesian(ylim = c(-2100, 3700)) +
	                  facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                    labs(title = "artificial contact material") + 
                    ylab(NULL) + xlab(NULL) +
	                  scale_colour_futurama() +
	                  theme_classic() +
                    theme(legend.position = "none") +
                    theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
	# combine both plots 
  Sq.bone <- Sq_bone.mean + Sq.plate.mean + plot_layout(width = c(3/6, 3/6), 
             guides = 'collect')  
  print(Sq.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Sq.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Sq.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## cow scapula & bone plate (with specific range per parameter) 
### Vmc
```{r}
  # plot cow scapula 
  Vmc_bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
                   mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                   y = "Vmc.diff.mean", colour = "Sample")) + 
                   geom_point(size = 4) +
                   coord_cartesian(ylim = c(-2, 2.7)) +
	                 facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                   labs(title = "natural contact material") + 
                   ylab("Vmc [µm³/µm²]") + xlab(NULL) +
	                 scale_colour_futurama() +
	                 theme_classic() +
                   theme(legend.position = "none") +
                   theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

   # plot bone plate 
   Vmc.plate.mean <- ggplot(data = mean.sample[grep("bone plate", 
                     mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                     y = "Vmc.diff.mean", colour = "Sample")) + 
	                   geom_point(size = 4) +
                     coord_cartesian(ylim = c(-2, 2.7)) +
	                   facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                     labs(title = "artificial contact material") + 
                     ylab(NULL) + xlab(NULL) +
	                   scale_colour_futurama() +
	                   theme_classic() +
                     theme(legend.position = "none") +
                     theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
	# combine both plots 
  Vmc.bone <- Vmc_bone.mean + Vmc.plate.mean + plot_layout(width = c(3/6, 3/6),                           guides = 'collect')  
  print(Vmc.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Vmc.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Vmc.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## cow scapula & bone plate (with specific range per parameter)
### HAsfc9
```{r}
  # plot cow scapula
  HAsfc9_bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "HAsfc9.diff.mean", colour = "Sample")) + 
                      geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.5, 10.5)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "natural contact material") + 
                      ylab("HAsfc9") + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot bone plate 
   HAsfc9.plate.mean <- ggplot(data = mean.sample[grep("bone plate", 
                        mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                        y = "HAsfc9.diff.mean", colour = "Sample")) + 
	                      geom_point(size = 4) +
                        coord_cartesian(ylim = c(-0.5, 10.5)) +
	                      facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                        labs(title = "artificial contact material") + 
                        ylab(NULL) + xlab(NULL) +
	                      scale_colour_futurama() +
	                      theme_classic() +
                        theme(legend.position = "none") +
                        theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  HAsfc9.bone <- HAsfc9_bone.mean + HAsfc9.plate.mean + plot_layout(width = c(3/6, 3/6),
                 guides = 'collect')  
  print(HAsfc9.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_HAsfc9.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = HAsfc9.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## cow scapula & bone plate (with specific range per parameter)
### epLsar
```{r}
  # plot cow scapula
  epLsar_bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "epLsar.diff.mean", colour = "Sample")) + 
                      geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.002, 0.004)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "natural contact material") + 
                      ylab("epLsar") + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
   # plot bone plate 
   epLsar.plate.mean <- ggplot(data = mean.sample[grep("bone plate", 
                        mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                        y = "epLsar.diff.mean", colour = "Sample")) + 
	                      geom_point(size = 4) +
                        coord_cartesian(ylim = c(-0.002, 0.004)) +
	                      facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                        labs(title = "artificial contact material") + 
                        ylab(NULL) + xlab(NULL) +
	                      scale_colour_futurama() +
	                      theme_classic() +
                        theme(legend.position = "none") +
                        theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  epLsar.bone <- epLsar_bone.mean + epLsar.plate.mean + plot_layout(width = c(3/6, 3/6),
                 guides = 'collect')  
  print(epLsar.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_epLsar.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = epLsar.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## cow scapula & bone plate (with specific range per parameter)
### Asfc
```{r}
  # plot cow scapula
  Asfc_bone.mean <- ggplot(data = mean.sample[grep("cow scapula", 
                    mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                    y = "Asfc.diff.mean", colour = "Sample")) + 
                    geom_point(size = 4) +
                    coord_cartesian(ylim = c(-16, 26)) +
	                  facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                    labs(title = "natural contact material") + 
                    ylab("Asfc") + xlab(NULL) +
	                  scale_colour_futurama() +
	                  theme_classic() +
                    theme(legend.position = "none") +
                    theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
 
   # plot bone plate 
   Asfc.plate.mean <- ggplot(data = mean.sample[grep("bone plate", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "Asfc.diff.mean", colour = "Sample")) + 
	                    geom_point(size = 4) +
                      coord_cartesian(ylim = c(-16, 26)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "artificial contact material") + 
                      ylab(NULL) + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  Asfc.bone <- Asfc_bone.mean + Asfc.plate.mean + plot_layout(width = c(3/6, 3/6), 
               guides = 'collect')  
  print(Asfc.bone)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Asfc.bone.mean", ".pdf")
	ggsave(filename = file_out, plot = Asfc.bone, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## pig skin & skin pad (with specific range per parameter)
### Sq
```{r}
  # plot pig skin 
  Sq_skin.mean <- ggplot(data = mean.sample[grep("pig skin", 
                  mean.sample[["Contact.material"]]), ], aes_string(x = "Sample",  
                  y = "Sq.diff.mean", colour = "Sample")) + 
                  geom_point(size = 4) +
                  coord_cartesian(ylim = c(-2100, 3700)) +
	                facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                  labs(title = "natural contact material") + 
                  ylab("Sq [nm]") + xlab(NULL) +
	                scale_colour_futurama() +
	                theme_classic() +
                  theme(legend.position = "none") +
                  theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot skin pad 
   Sq.pad.mean <- ggplot(data = mean.sample[grep("skin pad", 
                  mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                  y = "Sq.diff.mean", colour = "Sample")) + 
	                geom_point(size = 4) +
                  coord_cartesian(ylim = c(-2100, 3700)) +
	                facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                  labs(title = "artificial contact material") + 
                  ylab(NULL) + xlab(NULL) +
	                scale_colour_futurama() +
	                theme_classic() +
                  theme(legend.position = "none") +
                  theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  Sq.skin <- Sq_skin.mean + Sq.pad.mean + plot_layout(width = c(3/6, 3/6), 
             guides = 'collect')  
  print(Sq.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Sq.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Sq.skin, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## pig skin & skin pad (with specific range per parameter)
### Vmc
```{r}
  # plot pig skin 
  Vmc_skin.mean <- ggplot(data = mean.sample[grep("pig skin", 
                   mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                   y = "Vmc.diff.mean", colour = "Sample")) + 
                   geom_point(size = 4) +
                   coord_cartesian(ylim = c(-2, 2.7)) +
	                 facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                   labs(title = "natural contact material") + 
                   ylab("Vmc [µm³/µm²]") + xlab(NULL) +
	                 scale_colour_futurama() +
	                 theme_classic() +
                   theme(legend.position = "none") +
                   theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	
   # plot skin pad 
   Vmc.pad.mean <- ggplot(data = mean.sample[grep("skin pad", 
                   mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                   y = "Vmc.diff.mean", colour = "Sample")) + 
	                 geom_point(size = 4) +
                   coord_cartesian(ylim = c(-2, 2.7)) +
	                 facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                   labs(title = "artificial contact material") + 
                   ylab(NULL) + xlab(NULL) +
	                 scale_colour_futurama() +
	                 theme_classic() +
                   theme(legend.position = "none") +
                   theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  Vmc.skin <- Vmc_skin.mean + Vmc.pad.mean + plot_layout(width = c(3/6, 3/6), 
              guides = 'collect')  
  print(Vmc.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Vmc.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Vmc.skin, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## pig skin & skin pad (with specific range per parameter)
### HAsfc9
```{r}
  # plot pig skin 
  HAsfc9_skin.mean <- ggplot(data = mean.sample[grep("pig skin", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "HAsfc9.diff.mean", colour = "Sample")) + 
                      geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.5, 10.5)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "natural contact material") + 
                      ylab("HAsfc9") + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot skin pad 
   HAsfc9.pad.mean <- ggplot(data = mean.sample[grep("skin pad", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "HAsfc9.diff.mean", colour = "Sample")) + 
	                    geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.5, 10.5)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "artificial contact material") + 
                      ylab(NULL) + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  HAsfc9.skin <- HAsfc9_skin.mean + HAsfc9.pad.mean + plot_layout(width = c(3/6, 3/6),
                 guides = 'collect')  
  print(HAsfc9.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_HAsfc9.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = HAsfc9.skin, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```

## pig skin & skin pad (with specific range per parameter)
### epLsar
```{r}
  # plot pig skin 
  epLsar_skin.mean <- ggplot(data = mean.sample[grep("pig skin",  
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "epLsar.diff.mean", colour = "Sample")) + 
                      geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.002, 0.004)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "natural contact material") + 
                      ylab("epLsar") + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot skin pad 
   epLsar.pad.mean <- ggplot(data = mean.sample[grep("skin pad", 
                      mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                      y = "epLsar.diff.mean", colour = "Sample")) + 
	                    geom_point(size = 4) +
                      coord_cartesian(ylim = c(-0.002, 0.004)) +
	                    facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                      labs(title = "artificial contact material") + 
                      ylab(NULL) + xlab(NULL) +
	                    scale_colour_futurama() +
	                    theme_classic() +
                      theme(legend.position = "none") +
                      theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  epLsar.skin <- epLsar_skin.mean + epLsar.pad.mean + plot_layout(width = c(3/6, 3/6),
                 guides = 'collect')  
  print(epLsar.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_epLsar.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = epLsar.skin, path = dir_out, device = "pdf", 
	        width = 300, height = 170, units = "mm")


```

## pig skin & skin pad (with specific range per parameter)
### Asfc
```{r}
  # plot pig skin 
  Asfc_skin.mean <- ggplot(data = mean.sample[grep("pig skin", 
                    mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                    y = "Asfc.diff.mean", colour = "Sample")) + 
                    geom_point(size = 4) +
                    coord_cartesian(ylim = c(-16, 26)) +
	                  facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                    labs(title = "natural contact material") + 
                    ylab("Asfc") + xlab(NULL) +
	                  scale_colour_futurama() +
	                  theme_classic() +
                    theme(legend.position = "none") +
                    theme(axis.text.x = element_blank(),axis.ticks = element_blank())  
	            
   # plot skin pad 
   Asfc.pad.mean <- ggplot(data = mean.sample[grep("skin pad", 
                    mean.sample[["Contact.material"]]), ], aes_string(x = "Sample", 
                    y = "Asfc.diff.mean", colour = "Sample")) + 
	                  geom_point(size = 4) +
                    coord_cartesian(ylim = c(-16, 26)) +
	                  facet_wrap(Sample ~ Contact.material, nrow = 2) + 
                    labs(title = "artificial contact material") + 
                    ylab(NULL) + xlab(NULL) +
	                  scale_colour_futurama() +
	                  theme_classic() +
                    theme(legend.position = "none") +
                    theme(axis.text.x = element_blank(),axis.ticks = element_blank())  

	# combine both plots 
  Asfc.skin <- Asfc_skin.mean + Asfc.pad.mean + plot_layout(width = c(3/6, 3/6), 
               guides = 'collect')  
  print(Asfc.skin)

  # save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_Asfc.skin.mean", ".pdf")
	ggsave(filename = file_out, plot = Asfc.skin, path = dir_out, device = "pdf", 
	       width = 300, height = 170, units = "mm")


```


# Principal component analysis 
## PCA contact material - raw material seperated
### flint
```{r}
# filter the data for flint only
flint <- filter(imp_data, Raw.material == "flint") 

# remove rows with na values 
data_pca.flint <- na.omit(flint)
# use for the PCA only selected variables: Sq, SSK, Vmc, Isotropy, Mean density of furrows,
# Asfc, HAsfc9 
flint.pca <- prcomp(data_pca.flint[ , c(9:10, 27, 32:33, 40, 42)], scale. = TRUE) 
# convert the data into factor 
data_pca.flint[["Contact.material"]] <- factor(data_pca.flint[["Contact.material"]])

# using ggfortify
PCA.flint <- autoplot(flint.pca, data = data_pca.flint, colour = "Contact.material", 
             size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.repel = TRUE,  
             frame.alpha = 0) + 
             theme_classic() +
             scale_colour_futurama() 
         
print(PCA.flint)

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA.flint_contact", ".pdf")
ggsave(filename = file_out, plot = PCA.flint, path = dir_out, device = "pdf")

```


### lydite
```{r}
# filter the data for lydite only
lydite <- filter(imp_data, Raw.material == "lydite") 

# remove rows with na values 
data_pca.lydite <- na.omit(lydite)
# use for the PCA only selected variables: Sq, SSK, Vmc, Isotropy, Mean density of furrows,
# Asfc, HAsfc9 
lydite.pca <- prcomp(data_pca.lydite[ , c(9:10, 27, 32:33, 40, 42)], scale. = TRUE) 
# convert the data into factor 
data_pca.lydite[["Contact.material"]] <- factor(data_pca.lydite[["Contact.material"]])

# using ggfortify
PCA.lydite <- autoplot(lydite.pca, data = data_pca.lydite, colour = "Contact.material", 
              size = 2,
              loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
              loadings.label.colour = "black", 
              loadings.label.size  = 4, loadings.label.repel = TRUE,  
              frame.alpha = 0) + 
              theme_classic() +
              scale_colour_futurama() 
         
print(PCA.lydite)

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA.lydite_contact", ".pdf")
ggsave(filename = file_out, plot = PCA.lydite, path = dir_out, device = "pdf")

```


## PCA before and after - raw material seperated 
### flint
```{r}
# using ggfortify
PCA2_flint<- autoplot(flint.pca, data = data_pca.flint, colour = "Cycle", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, 
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.repel = TRUE,  
             frame = TRUE, frame.type = "convex", frame.colour = "Cycle", frame.alpha = 0) +
             theme_classic() +
             scale_colour_manual(values = custom.col6$col)
         
print(PCA2_flint)

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA.flint_cycle", ".pdf")
ggsave(filename = file_out, plot = PCA2_flint, path = dir_out, device = "pdf")
  
```


### lydite
```{r}
# using ggfortify
PCA2_lydite<- autoplot(lydite.pca, data = data_pca.lydite, colour = "Cycle", size = 2,
              loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, 
              loadings.label.colour = "black", 
              loadings.label.size  = 4, loadings.label.repel = TRUE,  
              frame = TRUE, frame.type = "convex", frame.colour = "Cycle", 
              frame.alpha = 0) +
              theme_classic() +
              scale_colour_manual(values = custom.col6$col)
         
print(PCA2_lydite)

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA.lydite_cycle", ".pdf")
ggsave(filename = file_out, plot = PCA2_lydite, path = dir_out, device = "pdf")
  
```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis_before.after/scripts/RStudioVersion.txt", n = 1)`.


---

# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)
```



---

END OF SCRIPT
